---
title: "Cell Activation Analysis Final Thesis"
output:
  pdf_document: default
date: "01 February 2019"
---

## Load Packages 

```{r loadpack, message=FALSE}
librarian::shelf(tidyverse, scales, DescTools, ggpubr, gmodels, 
                 knitr, kableExtra, broom, haven, readxl, RColorBrewer)
```

## read in cell activation data

```{r get_cell_act}
cell_act <- readRDS("cell_act_final.rds")

```


## Summary data of groups by resistance markers on cd8+ T cells -- not adjusted for cv

```{r group_act}
group_act <- cell_act %>% 
   group_by(group, visit) %>% 
   summarise(cd38_hladr_mean = mean(cd8_cd38_hladr_pct, na.rm = TRUE),
             cd38_hladr_sd = sd(cd8_cd38_hladr_pct, na.rm = TRUE),
             cd38_mean = mean(cd38, na.rm = TRUE),
             cd38_sd = sd(cd38, na.rm = TRUE),
             hla_dr_mean = mean(hla_dr, na.rm = TRUE),
             hla_dr_sd = sd(hla_dr, na.rm = TRUE))
kable(group_act)
```

## Graph of Difference among Reversion Groups for %CD38+ HLA-DR+ CD8 T-cells

```{r gg_cd38hla, message = FALSE, warning = FALSE}

my_comparisons <- list(c("none", "total"), c("total", "partial"))
gg_box_cd38hla <- cell_act %>% 
   mutate(visit = case_when(
      visit == 1 ~ "baseline",
      visit == 2 ~ "week 4",
      visit == 3 ~ "week 8",
      visit == 4 ~ "week 12"
   )) %>% 
   mutate(visit = factor(visit, levels = c("baseline", "week 4", "week 8", 
                                           "week 12"))) %>% 
   ggboxplot(x = "group", y = "cd8_cd38_hladr_pct",
             color = "group", 
             palette = "uchicago",
             title = "",
             xlab = "Reversion Group", 
             ylab = "%CD38+ HLA-DR+ CD8 T-cells",
             add = "jitter",
             legend = "none",
             ggtheme = theme_grey())
gg_box_cd38hla <- gg_box_cd38hla + stat_compare_means(comparisons = my_comparisons, 
                                    label = "p.format")
## Log graph for 2nd panel of graph

gg_box_log_cd38hla <- cell_act %>% 
   mutate(visit = case_when(
      visit == 1 ~ "baseline",
      visit == 2 ~ "week 4",
      visit == 3 ~ "week 8",
      visit == 4 ~ "week 12"
   )) %>% 
   mutate(visit = factor(visit, levels = c("baseline", "week 4", "week 8", 
                                           "week 12")),
          y_correct = cd8_cd38_hladr_pct/log) %>% 
   ggboxplot(x = "group", y = "y_correct",
             color = "group", 
             palette = "uchicago",
             title = "",
             xlab = "", 
             ylab = "",
             add = "jitter",
             legend = "none",
             ggtheme = theme_grey())
gg_box_log_cd38hla <- gg_box_log_cd38hla + stat_compare_means(comparisons = my_comparisons, 
                                    label = "p.format")

gg_box_CA_plots <- ggarrange(gg_box_cd38hla, gg_box_log_cd38hla,  
                  ncol = 2, nrow = 1,
                  labels = c("Uncorrected", "Corrected"),
                  vjust = 2.5)

gg_box_CA_plots

saveRDS(gg_box_CA_plots, file = "gg_box_CA_plots.rds")
```

## Correlation between Viral Load and Cell Activation

### Figure 

```{r corr_cv_cell_act}

gg_cv_cell_act <- cell_act %>% 
   filter_at(vars(c(cd38, log)), all_vars(!is.na(.))) %>% 
   ggscatter(x = "log", y = "cd38", 
          add = "reg.line", jitter = TRUE, conf.int = TRUE, 
          palette = "aaas",
          xlab = "Viral Load (log10 copies/mL)",
          ylab = "CD38 CD8+ T cells Count",  
          cor.coef = TRUE,
          ggtheme = theme_grey(), repel = TRUE, 
#          cor.coef.coord = c(4.5, 310),
          cor.coeff.args = list(method = "spearman"))
gg_cv_cell_act
   
## cell activation with CD4+

gg_cd4_cell_act <- cell_act %>% 
   filter_at(vars(c(cd38, cd4_total)), all_vars(!is.na(.))) %>% 
   ggscatter(y = "cd4_total", x = "log", 
          add = "reg.line", jitter = TRUE, conf.int = TRUE, 
          palette = "aaas",
          xlab = "Viral Load (log10 copies/mL)",
          ylab = "CD4+ T Cell Count / Cells per mL", 
          cor.coef = TRUE,
          ggtheme = theme_grey(), repel = TRUE, 
#          cor.coef.coord = c(0, 6.5),
          cor.coeff.args = list(method = "spearman"))
gg_cd4_cell_act

## join two plots together

cd38_plots <- ggarrange(gg_cv_cell_act, gg_cd4_cell_act,  
                  ncol = 2, nrow = 1,
                  labels = c("Cell Activation", "CD4+"))
cd38_plots

```

### Correlation of cd38 with viral load and cd4

```{r}
cor(cell_act[,c("log", "cd4_total", "cd38")], method = "spearman", 
    use = "complete.obs")
```

### Regression of cd38 against viral load

```{r cd38_reg}

cd38_vl <- lm(cd38 ~ log, data = cell_act)
summary(cd38_vl)

```

### Regression of cd38 against CD4+

```{r cd38_reg_cd4}

cd4_cd38 <- lm(cd4_total ~ cd38, data = cell_act)
summary(cd4_cd38)

```

## Relation of HLA-DR with CV and CD4+

```{r hla}
cor(cell_act[,c("log", "cd4_total", "hla_dr")], method = "spearman", 
    use = "complete.obs")

hla_vl <- lm(hla_dr ~ log, data = cell_act)
summary(hla_vl)

cd4_hla <- lm(cd4_total ~ hla_dr, data = cell_act)
summary(cd4_hla)

```

## Relation of Joint % on CD8+ with CV and CD4+

```{r joint}
cor(cell_act[,c("log", "cd4_total", "cd8_cd38_hladr_pct")], method = "spearman", 
    use = "complete.obs")

joint_vl <- lm(cd8_cd38_hladr_pct ~ log, data = cell_act)
summary(joint_vl)

cd4_joint <- lm(cd4_total ~ cd8_cd38_hladr_pct, data = cell_act)
summary(cd4_joint)

```


## Models of CD4 Decrease in Relation to Viral Load and/or CD38 increase

```{r models}
# Create data frame and remove cases with NA in 
res_pacs <- cell_act %>% 
   filter(!is.na(cd4_total) & !is.na(log) & !is.na(cd38))

## Model 1 - just viral load
mod1 <- lm(cd4_total ~ log, data = res_pacs)

## Model 2 - just cd38

mod2 <-  lm(cd4_total ~ cd38, data = res_pacs)

## Model 3 - both viral load and cd 38

mod3 <- lm(cd4_total ~ log + cd38, data = res_pacs)

## Model 4 -- both factors + interaction

mod4 <- lm(cd4_total ~ log + cd38 + log:cd38, data = res_pacs)

## Model 5 - joint factors 

mod5 <-  lm(cd4_total ~ cd8_cd38_hladr_pct, data = res_pacs)

## Model 6 -- joint + viral load

mod6 <- lm(cd4_total ~ log + cd8_cd38_hladr_pct, data = res_pacs)
#
# group together the models
### coefficients data frame
regcoef <- tidy(mod1) %>% 
   add_column(model = "logcv", .before = "term" )
regcoef <- rbind(regcoef, tidy(mod2) %>% 
   add_column(model = "cd38", .before = "term"))
regcoef <- rbind(regcoef, tidy(mod3) %>% 
   add_column(model = "both", .before = "term"))
regcoef <- rbind(regcoef, tidy(mod4) %>% 
   add_column(model = "inter", .before = "term"))
regcoef <- rbind(regcoef, tidy(mod5) %>% 
   add_column(model = "cd38+HLA-DR+", .before = "term"))
regcoef <- rbind(regcoef, tidy(mod6) %>% 
   add_column(model = "cd38+HLA-DR+/Log", .before = "term"))

### summary data frame
regsumm <- glance(mod1) %>% 
   add_column(model = "logcv", .before = "r.squared" )
regsumm <- rbind(regsumm, glance(mod2) %>% 
   add_column(model = "cd38", .before = "r.squared"))
regsumm <- rbind(regsumm, glance(mod3) %>% 
   add_column(model = "both", .before = "r.squared"))
regsumm <- rbind(regsumm, glance(mod4) %>% 
   add_column(model = "inter", .before = "r.squared"))
regsumm <- rbind(regsumm, glance(mod5) %>% 
   add_column(model = "cd38+HLA-DR+", .before = "r.squared"))
regsumm <- rbind(regsumm, glance(mod6) %>% 
   add_column(model = "cd38+HLA-DR+/Log", .before = "r.squared"))
regsumm <- regsumm %>% select(1:9) # reduce num of measures

kable(regcoef)
kable(regsumm)

```

## Control Cell Activation by Viral Load

Calculate a new index of cell activation (`cd8_cd38_hladr_pct`) divided by log viral load. Measure levels of this per group and compare to cd4+ levels by group

```{r cell_act_control}
cell_act <- cell_act %>% 
   mutate(act_cv = cd8_cd38_hladr_pct/log)
Desc(cell_act$act_cv, plotit = TRUE)
Desc(act_cv ~ group, data = cell_act, plotit = TRUE)
```

## Graph of act_cv by group

```{r gg_actcv, message = FALSE, warning = FALSE}
gg_actcv <- cell_act %>% 
   ggboxplot(x = "group", y = "act_cv", 
             color = "group", 
             palette = "aaas",
             title = "Cellular Activation by Reversion Group",
             subtitle = "Controlled for by Viral Load",
             xlab = "Reversion Group", 
             ylab = "Activation/Viral Load Index",
             add = "jitter",
             legend = "none",
             ggtheme = theme_grey())
my_comparisons <- list(c("none", "total"), c("total", "partial"))
gg_actcv + stat_compare_means(comparisons = my_comparisons, 
                                    label = "p.format")
saveRDS(gg_actcv, file = "gg_box_actcv.rds")   

```

```{r gg_actcv_cd4, message = FALSE, warning = FALSE}
gg_actcv_cd4 <- cell_act %>% 
   ggscatter(y = "cd4_total", x = "act_cv", 
             color = "group", 
             palette = "aaas",
             add = "reg.line", 
             title = "CD4+ T-Cell Levels vs. Cellular Activation",
             subtitle = "Controlled for by Viral Load",
             xlab = "Activation/Viral Load Index", 
             ylab = "CD4+ T-cell count (cells/mL)",
             cor.coef.coord = c(4.5, 310),
             cor.coeff.args = list(method = "spearman"),
             legend = "bottom",
             ggtheme = theme_grey())
gg_actcv_cd4

```

## Regressions by Group of CD4 vs. Cellular Activation Index

```{r reg_cd4_cell_act_ind}
# all resistants patients
reg_all <- lm(cd4_total ~ act_cv, data = cell_act)
summary(reg_all)

# group = none
reg_none <- lm(cd4_total ~ act_cv, data = cell_act, 
               subset = (cell_act$group == "none"))
summary(reg_none)

# group = total
reg_total <- lm(cd4_total ~ act_cv, data = cell_act, 
               subset = (cell_act$group == "total"))
summary(reg_total)

# group = partial
reg_part <- lm(cd4_total ~ act_cv, data = cell_act, 
               subset = (cell_act$group == "partial"))
summary(reg_part)

reg_activ <- tidy(reg_all) %>% 
   add_column(model = "all resistant", .before = "term")

reg_activ <- rbind(reg_activ, tidy(reg_none) %>% 
   add_column(model = "no reversion", .before = "term"))

reg_activ <- rbind(reg_activ, tidy(reg_total) %>% 
   add_column(model = "total reversion", .before = "term"))

reg_activ <- rbind(reg_activ, tidy(reg_part) %>% 
   add_column(model = "partial reversion", .before = "term"))

regsumm_activ <- glance(reg_all) %>% 
   add_column(model = "all resistant", .before = "r.squared" )

regsumm_activ <- rbind(regsumm_activ, glance(reg_none) %>% 
   add_column(model = "no reversion", .before = "r.squared"))

regsumm_activ <- rbind(regsumm_activ, glance(reg_total) %>% 
   add_column(model = "total reversion", .before = "r.squared"))

regsumm_activ <- rbind(regsumm_activ, glance(reg_part) %>% 
   add_column(model = "partial reversion", .before = "r.squared"))

kable(reg_activ)
regsumm_activ %>% 
   select(-c(sigma, df, logLik, BIC, deviance, df.residual)) %>% 
   kable()
```

## Cellular activation by week by group

```{r cell_act_wk_group}
options(scipen = 5)

group_wk_act <- cell_act %>% 
    mutate(visit = case_when(
      visit == 1 ~ "baseline",
      visit == 2 ~ "week 4",
      visit == 3 ~ "week 8",
      visit == 4 ~ "week 12"
    )) %>% 
   select(subject:visit, group, log, cd8_cd38_hladr_pct) %>% 
   mutate(act_cv = cd8_cd38_hladr_pct/log) %>%  # control for carga viral
   mutate(visit = factor(visit, levels = c("baseline", "week 4", "week 8", 
                                          "week 12")))  
group_wk_act_summ <- group_wk_act %>% 
   group_by(group, visit) %>% 
   summarise(cd38_hladr_mean = mean(cd8_cd38_hladr_pct, na.rm = TRUE),
             cd38_hladr_sd = sd(cd8_cd38_hladr_pct, na.rm = TRUE),
             corr_mean = mean(act_cv, na.rm = TRUE),  # corrected for cv
             corr_hladr_sd = sd(act_cv, na.rm = TRUE)) # corrected for cv
   
group_wk_act_summ %>% 
   select(group, visit, cd38_hladr_mean, corr_mean) %>%
   filter(group != "partial") %>% 
   kable(booktabs = TRUE, col.names = c("Group", "Visit", "Raw Mean", 
                                        "Corrected Mean"), 
   caption = "Mean Activation Evolution by Group and Visit")  
```

## Graphs of Evolution of Activation based on Corrected Values

```{r gg_evolution_act, warning = FALSE, message = FALSE}
# my_comparisons <- list(c("baseline", "week 12"))
gg_group_wk_act <- group_wk_act %>% 
   filter(group != "partial") %>% #(make it total vs. none)
   ggerrorplot(., x = "visit", y = "act_cv", 
               desc_stat = "mean_se",
               color = "black",
               palette = "uchicago",
               title = "Cell Activation Evolution in None and Total Reversion Groups",
               subtitle = "Activation Measure Corrected for Viral Load",
               xlab = "Visit",
               ylab = "Activation/Viral Load Index",
               add = "jitter",
               add.params = list(color = "darkgray"),
               facet.by = "group",
#               panel.labs = list(rev = c("No Reversion", "Total Reversion")),
               ggtheme = theme_gray())
#gg_group_wk_act <- gg_group_wk_act + stat_compare_means(comparisons = 
#                                                           my_comparisons)
gg_group_wk_act
```

## Graphs of Evolution of Activation based on RAw Values

```{r gg_evolution_act_raw, warning = FALSE, message = FALSE}
# my_comparisons <- list(c("baseline", "week 12"))
gg_group_wk_act_r <- group_wk_act %>% 
   filter(group != "partial") %>% #(make it total vs. none)
   ggerrorplot(., x = "visit", y = "cd8_cd38_hladr_pct", 
               desc_stat = "mean_se",
               color = "black",
               palette = "uchicago",
               title = "Cell Activation Evolution in None and Total Reversion Groups",
#               subtitle = "Activation Measure Corrected for Viral Load",
               xlab = "Visit",
               ylab = "CD8 + with CD38 and HLA-DR",
               add = "jitter",
               add.params = list(color = "darkgray"),
               facet.by = "group",
#               panel.labs = list(rev = c("No Reversion", "Total Reversion")),
               ggtheme = theme_gray())
#gg_group_wk_act <- gg_group_wk_act + stat_compare_means(comparisons = 
#                                                           my_comparisons)
gg_group_wk_act_r
```


## statistical difference between None and Total in terms of Difference Week 12 - baseline

```{r stat_diff_wk12_bl}
pacs <- unique(group_wk_act$subject)

diffs <- tibble(subject = character(),
                group = factor(),
                raw_diff = numeric(),
                corr_diff = numeric())

for (i in 1:length(pacs)) {
   raw_diff <- group_wk_act$cd8_cd38_hladr_pct[group_wk_act$subject == pacs[i] &
                                              group_wk_act$visit == "week 12"] - 
               group_wk_act$cd8_cd38_hladr_pct[group_wk_act$subject == pacs[i] &
                                              group_wk_act$visit == "baseline"]
   corr_diff <- group_wk_act$act_cv[group_wk_act$subject == pacs[i] &
                                              group_wk_act$visit == "week 12"] - 
               group_wk_act$act_cv[group_wk_act$subject == pacs[i] &
                                              group_wk_act$visit == "baseline"]
   diffs <- add_row(diffs, subject = pacs[i],
                    group = group_wk_act$group[group_wk_act$subject == pacs[i] &
                                                  group_wk_act$visit == "week 12"],
                    raw_diff = raw_diff,
                    corr_diff = corr_diff)
} 

diffs <- diffs %>% 
   filter(group != "partial") %>% 
   droplevels(.)

diffs

# Do the comparison raw data

Desc(raw_diff ~ group, data = diffs)
Desc(corr_diff ~ group, data = diffs)
```

