---
title: "Comparison of Regression and Survival"
author: "James R. Hunter"
date: "01 August 2018"
output: 
  pdf_document:
    fig_height: 7
    fig_width: 7
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
version: 4
---

```{r, setup, include=FALSE}
require(mosaic)   # Load additional packages here
suppressMessages(library(tidyverse))
suppressPackageStartupMessages(library(DescTools))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(RColorBrewer))
suppressPackageStartupMessages(library(seqinr))
suppressPackageStartupMessages(library(ggpubr))
suppressPackageStartupMessages(library(kableExtra))
suppressPackageStartupMessages(library(Biostrings))
library(survival)
library(survminer)
library(stringr)
options(scipen = 1)
  
knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  echo = FALSE,   # don't show programming
  size="small")   # slightly smaller font for code

rev_groups <- readRDS("rev_groups.rds")
pacs_resist <- rev_groups  %>% 
   filter(rev_group != 0)
pacs_resist <- as.numeric(pacs_resist$n)
rev_groups <- rev_groups %>% 
   filter(rev_group != 0) %>% 
   select(rev_group)
load("prpbm2.RData")
load("trpbm2.RData")
```

## Regression -- PI 46 and 90 Reversion Time Plots

### Load prpbmc mutations

```{r loadprmuts, echo = FALSE}
prpbmuts_rev <- read_rds("prpbmuts_rev.rds")
```

### Create Function to Calculate regression and make plot

```{r fig5funct, echo = FALSE}
drawPRmut <- function(pimut){
# Calculate regression equation
   fit <- lm(num ~ wk, data = prpbmuts_rev[prpbmuts_rev$type == pimut,])
   intercept <- unname(format(coef(fit)[1], digits = 3))
   slope <- unname(format(coef(fit)[2], digits = 3))
   rsq <- unname(format(summary(fit)$r.squared, digits = 3))
   eqn <- as.character(as.expression(
      substitute(italic(y) == a + b * italic(x) * "," ~~ italic(r)^2 ~ "=" ~ r2,
        list(a = intercept,
             b = slope,
             r2 = rsq   ))))
# Draw graph
   title <- paste(pimut)
   g1 <- ggplot(prpbmuts_rev[prpbmuts_rev$type == pimut,], aes(x = wk, y = num)) 
   g1 <- g1 + ggtitle(title)
   g1 <- g1 + xlab("Week") + ylab("Patients with Mutation")
   g1 <- g1 + geom_point()
   g1 <- g1 + geom_smooth(method = "lm")
   g1 <- g1 + scale_y_continuous(breaks = seq(from = 0, to = 40, by = 5))
   g1 <- g1 + scale_x_continuous(breaks = c(4, 8, 12)) 
   g1 <- g1 + annotate("text", x = 5, y = 2.5, label = eqn, parse = TRUE)
}
```

### Draw pr46 and pr90

```{r draw4690, echo = FALSE}
gr46 <- drawPRmut("pr46")
gr90 <- drawPRmut("pr90")
fig5 <- ggarrange(gr46, gr90, ncol = 1, nrow = 2, align = "v")
fig5

```

## Survival Version of Protease Inhibitor Mutations


```{r protease}
mutnamespr <- c("pr10", "pr20", "pr24", "pr30", "pr32", "pr36", "pr46", "pr47",
              "pr48", "pr50", "pr53", "pr54", "pr63", "pr71", "pr73", "pr77",
              "pr82", "pr84", "pr88", "pr89", "pr90", "pr93")
prmuts <- prpbm2 %>% 
      filter(type %in% mutnamespr & n %in% pacs_resist) %>% 
      mutate(wt = ifelse(value == 1, 0, 1)) 
surv_pr <- Surv(time = prmuts$wk, event = prmuts$wt)
fit_all_pr <- survfit(surv_pr ~ 1, data = prmuts)
fit_each_pr <- survfit(surv_pr ~ type, data = prmuts)
#summary(fit_all_pr)
summary(fit_each_pr)
title <- paste0("Conversion to Wild Type for Protease Codons")   
ggsurvplot(fit_all_pr, data = prmuts,
     palette = "aaas",
     break.x.by = 4,
     legend.title = "Codon",
#     pval = TRUE,
     xlab = "Weeks",
     ylab = "Mutation Retention Probability",
     title = title,
     ggtheme = theme_gray())

```

### Show Contrast between Best and Worst Codons (89 & 63)

```{r bestworstpr}
title <- paste0("Conversion to Wild Type for Codons pr63 and pr89") 
bestworst <- prpbm2 %>% 
      filter(type %in% c("pr63", "pr89") & n %in% pacs_resist) %>% 
      mutate(wt = ifelse(value == 1, 0, 1)) 
surv_pr_bw <- Surv(time = bestworst$wk, event = bestworst$wt)
fit_pr_bw <- survfit(surv_pr_bw ~ type, data = bestworst)
ggsurvplot(fit_pr_bw, data = bestworst,
     palette = "aaas",
     break.x.by = 4,
     legend.title = "Codon",
     pval = TRUE,
     xlab = "Weeks",
     ylab = "Mutation Retention Probability",
     title = title,
     ggtheme = theme_gray())
```

### Show contrast between pr46 and pr90 in Survival Analysis

```{r pr4690_surv}
title <- paste0("Conversion to Wild Type for Codons pr46 and pr90") 
pr4690 <- prpbm2 %>% 
      filter(type %in% c("pr46", "pr90") & n %in% pacs_resist) %>% 
      mutate(wt = ifelse(value == 1, 0, 1)) 
surv_pr_bw <- Surv(time = pr4690$wk, event = pr4690$wt)
fit_pr_bw <- survfit(surv_pr_bw ~ type, data = pr4690)
ggsurvplot(fit_pr_bw, data = pr4690,
     palette = "aaas",
     break.x.by = 4,
     legend.title = "Codon",
     pval = TRUE,
     xlab = "Weeks",
     ylab = "Mutation Retention Probability",
     title = title,
     ggtheme = theme_gray())
```

## Regression NRTIs -- Compare tr184 and TAMS

Using the data from Table 7 in the paper

```{r mutsxls, echo = FALSE}
trpbmuts_rev <- read_rds("trpbmuts_rev.rds")
trwksto0 <- read_rds("trweeksto0.rds")
nrti_muts <- c("tr41",  "tr210", "tr215", "tr67",  "tr70",  "tr219", "tr184")
nnrti_muts <- c("tr103", "tr181", "tr190")
tam1 <- c("tr41", "tr210", "tr215")
tam2 <- c("tr67", "tr70", "tr219")
nrtis <- trwksto0 %>%
   filter(Mutation %in% nrti_muts) %>% 
   mutate(ARVType = case_when(
      Mutation %in% tam1 ~ "NRTI - TAM 1",
      Mutation %in% tam2 ~ "NRTI - TAM 2",
      Mutation == "tr184" ~ "NRTI - tr184"
   ))

# Graph of weeks to 0
ggwksto0 <- ggscatter(nrtis, x = "ARVType", y = "Weeks",
          color = "ARVType",
          palette = "aaas",
          mean.point = TRUE,
          title = "Weeks to 0 Patients by NRTI Mutation Type",
          xlab = "Mutation Type",
          ylab = "Weeks to 0 Patients",
          ggtheme = theme_gray()) + 
          ylim(10, 35)

# ggwksto0

## prepare data to include in Graph in table
wkssumm <- nrtis %>% 
   group_by(ARVType) %>% 
   summarize(n = n(), avgweek = round(mean(Weeks), 3), 
             avgR2 = round(mean(r2), 3)) 
body_style <- tbody_style(size = 10)
tab_wkssumm <- ggtexttable(wkssumm, rows = NULL, 
                           theme = ttheme(base_style = "mBlackWhite")) 

## Join with graph
gg_wkssumm <- ggwksto0 +
   annotation_custom(ggplotGrob(tab_wkssumm), xmin = 0.9, xmax = 2.5,
                     ymin = 25)
gg_wkssumm

```

### Regression -- Figures with all NRTIs

#### Divide mutations into NRTI and NNRTI

```{r dividetrs, echo = FALSE}
nrti_muts <- c("tr41",  "tr210", "tr215", "tr67",  "tr70",  "tr219", "tr184")
nnrti_muts <- c("tr103", "tr181", "tr190")

trpbmuts <- read_rds("trpbmuts_rev.rds")
nrti <- trpbmuts %>% 
   filter(type %in% nrti_muts) %>% 
   mutate(mut_title = case_when(
      type %in% c("tr41", "tr210", "tr215") ~ 
         paste0(type, " - ", "NRTI – TAM 1"),
      type == "tr184" ~ 
         paste0(type, " - ", "NRTI - NAM"),
      TRUE ~ paste(type, "-", "NRTI – TAM 2"))) 

# set up df for all the TAMs for a graph
allTAMS <- nrti %>% 
   filter(type != "tr184") %>% 
   mutate(type = "allTAMs")
   
nnrti <- trpbmuts %>% 
   filter(type %in% nnrti_muts)
```

### Create Functions to Calculate regression

```{r regfunction}
drawTRmut <- function(trmut){
# Calculate regression equation
   fit <- lm(num ~ wk, data = trpbmuts[trpbmuts$type == trmut,])
   intercept <- unname(format(coef(fit)[1], digits = 3))
   slope <- unname(format(coef(fit)[2], digits = 3))
   rsq <- unname(format(summary(fit)$r.squared, digits = 3))
   eqn <- as.character(as.expression(
      substitute(italic(y) == a + b * italic(x) * "," ~~ italic(r)^2 ~ "=" ~ r2,
        list(a = intercept,
             b = slope,
             r2 = rsq
   ))))

# Draw graph
   title <- trmut
   g1 <- ggplot(nrti[nrti$type == trmut,], aes(x = wk, y = num)) 
   g1 <- g1 + ggtitle(title)
   g1 <- g1 + xlab("Week") + ylab("Patients with Mutation")
   g1 <- g1 + geom_point()
   g1 <- g1 + geom_smooth(method = "lm")
   g1 <- g1 + scale_y_continuous(breaks = seq(from = 0, to = 40, by = 5))
   g1 <- g1 + scale_x_continuous(breaks = c(4, 8, 12)) 
   g1 <- g1 + annotate("text", x = 5, y = 1, label = eqn, parse = TRUE)
}
```


## Function for all the AllTams Data
```{r alltamfunction}
drawalltam <- function(trmut){
# Calculate regression equation
   fit <- lm(num ~ wk, data = allTAMS[allTAMS$type == trmut,])
   intercept <- unname(format(coef(fit)[1], digits = 3))
   slope <- unname(format(coef(fit)[2], digits = 3))
   rsq <- unname(format(summary(fit)$r.squared, digits = 3))
   eqn <- as.character(as.expression(
      substitute(italic(y) == a + b * italic(x) * "," ~~ italic(r)^2 ~ "=" ~ r2,
        list(a = intercept,
             b = slope,
             r2 = rsq
   ))))

# Draw graph
   title <- trmut
   g1 <- ggplot(allTAMS[allTAMS$type == trmut,], aes(x = wk, y = num)) 
   g1 <- g1 + ggtitle(title)
   g1 <- g1 + xlab("Week") + ylab("Patients with Mutation")
   g1 <- g1 + geom_point()
   g1 <- g1 + geom_smooth(method = "lm")
   g1 <- g1 + scale_y_continuous(breaks = seq(from = 0, to = 40, by = 5))
   g1 <- g1 + scale_x_continuous(breaks = c(4, 8, 12)) 
   g1 <- g1 + annotate("text", x = 5, y = 1, label = eqn, parse = TRUE)
}
```

### Draw RT Mutation Projections

```{r drawfig5}
fig7a <- drawTRmut(nrti_muts[1])
fig7b <- drawTRmut(nrti_muts[2])
fig7c <- drawTRmut(nrti_muts[3])
fig7d <- drawTRmut(nrti_muts[4])
fig7e <- drawTRmut(nrti_muts[5])
fig7f <- drawTRmut(nrti_muts[6])
fig7g <- drawTRmut(nrti_muts[7])
fig7h <- drawalltam("allTAMs")

```

## Figure 8 -- NNRTIs 

## Function for NNRTI Data
```{r nnrtifunction}
draw_nnrti <- function(trmut){
# Calculate regression equation
   fit <- lm(num ~ wk, data = nnrti[nnrti$type == trmut,])
   intercept <- unname(format(coef(fit)[1], digits = 3))
   slope <- unname(format(coef(fit)[2], digits = 3))
   rsq <- unname(format(summary(fit)$r.squared, digits = 3))
   eqn <- as.character(as.expression(
      substitute(italic(y) == a + b * italic(x) * "," ~~ italic(r)^2 ~ "=" ~ r2,
        list(a = intercept,
             b = slope,
             r2 = rsq
   ))))

# Draw graph
   title <- trmut
   g1 <- ggplot(nnrti[nnrti$type == trmut,], aes(x = wk, y = num)) 
   g1 <- g1 + ggtitle(title)
   g1 <- g1 + xlab("Week") + ylab("Patients with Mutation")
   g1 <- g1 + geom_point()
   g1 <- g1 + geom_smooth(method = "lm")
   g1 <- g1 + scale_y_continuous(breaks = seq(from = 0, to = 40, by = 5))
   g1 <- g1 + scale_x_continuous(breaks = c(4, 8, 12)) 
   g1 <- g1 + annotate("text", x = 5, y = 1, label = eqn, parse = TRUE)
}
```

### Draw NNRTI's

```{r nnrtigr, echo = FALSE, message = FALSE, warning = FALSE}
fig6a <- draw_nnrti(nnrti_muts[1])
fig6b <- draw_nnrti(nnrti_muts[2])
fig6c <- draw_nnrti(nnrti_muts[3])
```

## Put NNRTI's in single graph

```{r combine_nnrti, echo = FALSE}
fignnrti <- ggarrange(fig6a, fig6b, fig6c, ncol = 3, nrow = 1)
fignnrti
```

## Survival -- NRTIs -- Compare TAMS and TAMS with M184V

```{r nrti}

tam1 <- c("tr41", "tr210", "tr215")
tam2 <- c("tr67", "tr70", "tr219")
all_tams <- c("tr41", "tr210", "tr215", "tr67", "tr70", "tr219")
nrti <- c("tr41", "tr210", "tr215", "tr67", "tr70", "tr219", "tr184")

```

### TAM1 Internal Differences

```{r tam1}
   name <- "TAM1"
   tam1case <- trpbm2 %>% 
      filter(type %in% tam1) %>% 
      mutate(wt = ifelse(value == 1, 0, 1))
   surv_tam1 <- Surv(time = tam1case$wk, event = tam1case$wt)
   fit_tam1 <- survfit(surv_tam1 ~ type, data = tam1case)
   title <- paste0("Conversion to Wild Type for ", name, " Codons")   
#   summary(fit_tam1)
   ggsurvplot(fit_tam1, data = tam1case,
        palette = "aaas",
        break.x.by = 4,
        legend.title = "Codon",
        pval = TRUE,
        xlab = "Weeks",
        ylab = "Mutation Retention Probability",
        title = title,
        ggtheme = theme_gray())
```

### TAM2 Internal Differences

```{r tam2}
   name <- "TAM2"
   tam2case <- trpbm2 %>% 
      filter(type %in% tam2) %>% 
      mutate(wt = ifelse(value == 1, 0, 1))
   surv_tam2 <- Surv(time = tam2case$wk, event = tam2case$wt)
   fit_tam2 <- survfit(surv_tam2 ~ type, data = tam2case)
   title <- paste0("Conversion to Wild Type for ", name, " Codons")   
#   summary(fit_tam2)
   ggsurvplot(fit_tam2, data = tam2case,
        palette = "aaas",
        break.x.by = 4,
        legend.title = "Codon",
        pval = TRUE,
        xlab = "Weeks",
        ylab = "Mutation Retention Probability",
        title = title,
        ggtheme = theme_gray())
```

### TAM1 vs. TAM2

```{r tam12}
   name <- "TAM"
   tam12case <- trpbm2 %>% 
      filter(type %in% all_tams) %>% 
      mutate(wt = ifelse(value == 1, 0, 1)) %>% 
      mutate(tam_type = ifelse(type %in% tam1, "TAM1", "TAM2"))
   surv_tam12 <- Surv(time = tam12case$wk, event = tam12case$wt)
   fit_tam12 <- survfit(surv_tam12 ~ tam_type, data = tam12case)
   title <- paste0("Conversion to Wild Type for ", name, " Codons")   
#   summary(fit_tam12)
   ggsurvplot(fit_tam12, data = tam12case,
        palette = "aaas",
        break.x.by = 4,
        legend.title = "Codon",
        pval = TRUE,
        xlab = "Weeks",
        ylab = "Mutation Retention Probability",
        title = title,
        ggtheme = theme_gray())

```

### All TAMs Taken as a Group

```{r all_TAMS}
name <- "All TAM"
   alltamcase <- trpbm2 %>% 
      filter(type %in% all_tams) %>% 
      mutate(wt = ifelse(value == 1, 0, 1)) 
   surv_alltam <- Surv(time = alltamcase$wk, event = alltamcase$wt)
   fit_alltam <- survfit(surv_alltam ~ 1, data = alltamcase)
   title <- paste0("Conversion to Wild Type for ", name, " Codons")   
   summary(fit_alltam)
   ggsurvplot(fit_alltam, data = alltamcase,
        palette = "aaas",
        break.x.by = 4,
        legend.title = "Codon",
        pval = TRUE,
        xlab = "Weeks",
        ylab = "Mutation Retention Probability",
        title = title,
        ggtheme = theme_gray())

```

### TAMS vs tr184

```{r TAMS184}
name <- "NRTI"
   tam184case <- trpbm2 %>% 
      filter(type %in% nrti) %>% 
      mutate(wt = ifelse(value == 1, 0, 1)) %>% 
      mutate(type = ifelse(type == "tr184", "tr184", "TAM"))
   surv_tam184 <- Surv(time = tam184case$wk, event = tam184case$wt)
   fit_tam184 <- survfit(surv_tam184 ~ type, data = tam184case)
   title <- paste0("Conversion to Wild Type for ", name, " Codons")   
#   summary(fit_tam184)
   ggsurvplot(fit_tam184, data = tam184case,
        palette = "aaas",
        break.x.by = 4,
        legend.title = "Codon",
        pval = TRUE,
        xlab = "Weeks",
        ylab = "Mutation Retention Probability",
        title = title,
        ggtheme = theme_gray())
```

### All NRTI Codons as a Group

```{r all_NRTI}
name <- "All NRTI"
   allcase <- trpbm2 %>% 
      filter(type %in% nrti) %>% 
      mutate(wt = ifelse(value == 1, 0, 1)) 
   surv_allnrti <- Surv(time = allcase$wk, event = allcase$wt)
   fit_allnrti <- survfit(surv_allnrti ~ 1, data = allcase)
   title <- paste0("Conversion to Wild Type for ", name, " Codons")   
#   summary(fit_allnrti)
   ggsurvplot(fit_allnrti, data = allcase,
        palette = "aaas",
        break.x.by = 4,
        legend.title = "Codon",
        pval = TRUE,
        xlab = "Weeks",
        ylab = "Mutation Retention Probability",
        title = title,
        ggtheme = theme_gray())

```

## Survival -- NNRTIs

1. Use the 3 NNRTI codons: 103, 181, 190
2. Modify the value as above


```{r nnrti}
nnrti_codons <- c("tr103", "tr181", "tr190")
nnrti <- trpbm2 %>% 
   filter(type %in% nnrti_codons) %>% 
   mutate(wt = ifelse(value == 1, 0, 1))
survnnrti <- Surv(time = nnrti$wk, event = nnrti$wt)
fitnnrti <- survfit(survnnrti ~ type, data = nnrti)
# summary(fitnnrti)
ggsurvplot(fitnnrti, data = nnrti,
           palette = "aaas",
           break.x.by = 4,
           legend.title = "Codon",
           xlab = "Weeks",
           ylab = "Mutation Retention Probability",
           title = "Conversion to Wild Type for NNRTI Codons",
           ggtheme = theme_gray())
```

## Positive Slope PR Codons -- 77 & 88

Show their results in survival analysis

```{r surv7788}
mutnamespr <- c("pr77", "pr88")
prmuts <- prpbm2 %>% 
       filter(type %in% mutnamespr & n %in% pacs_resist) %>% 
       mutate(wt = ifelse(value == 1, 0, 1))  # wt = wild type
surv_pr <- Surv(time = prmuts$wk, event = prmuts$wt)
fit_each_pr <- survfit(surv_pr ~ type, data = prmuts)
summary(fit_each_pr)

prmuts %>% 
   filter(type == "pr77") %>% 
   arrange(type, wk)
prmuts %>% 
   group_by(type, wk) %>% 
   summarise(wktots = sum(wt))
## Graph
title <- paste0("Reversion to Wild Type for Protease Codons 77 & 88")   
ggsurv7788 <- ggsurvplot(fit_each_pr, data = prmuts,
     palette = "aaas",
     break.x.by = 4,
     legend.title = "Codon",
#     pval = TRUE,
     cumevents = TRUE,
     xlab = "Weeks",
     ylab = "Mutation Retention Probability",
     title = title,
     ggtheme = theme_gray())
ggsurv7788
```

### Draw pr77 and pr88 as regression

```{r drawreg7788, echo = FALSE, }
wksto0_pr <- readRDS("wksto0_pr_full.rds")
edit_out_pr <- wksto0_pr$Mutation[wksto0_pr$Weeks == Inf]
resist_pac <- rev_groups %>% 
   filter(rev_group != 0) %>% 
   mutate(pac = n)
   select(pac)
resist_pac <- resist_pac$pac
prpbmuts <- prpbm2 %>% 
   filter(n %in% resist_pac) %>% 
   filter(value > 0) %>% 
   group_by(wk, type) %>% 
   summarize(num = n())
prpbmuts <- prpbmuts %>% filter(!(type %in% edit_out_pr))
prmutnames <- unique(prpbmuts$type)
wksto0 <- data_frame(Mutation = NA, Weeks = NA, int = NA, beta = NA, r2 = NA)
# Recalculate regression
# loop over mutations
for (i in 1:length(prmutnames)){
# Calculate regression equation
   fit <- lm(num ~ wk, data = prpbmuts[prpbmuts$type == prmutnames[i],])
# Put week to 0 patients calculation in wk0 data frame
   intercept <- fit$coef[1]
   slope <- fit$coef[2]
# algebraic solution
   wks <- (-intercept/slope)
   wksto0 <- rbind(wksto0, data_frame(Mutation = as.character(prmutnames[i]), 
                                      Weeks = wks,
                                      int = intercept,
                                      beta = slope,
                                      r2 = summary(fit)$r.squared))
}
prwksto0 <- wksto0[-1,]

drawPRmut <- function(prmut){
   mut <- prwksto0 %>% 
      filter(Mutation == prmut) 
   intercept <- unname(format(mut$int[1], digits = 3))
   slope <- unname(format(mut$beta[1], digits = 3))
   rsq <- unname(format(mut$r2[1], digits = 3))
   eqn <- as.character(as.expression(
      substitute(italic(y) == a + b * italic(x) * "," ~~ italic(r)^2 ~ "=" ~ r2,
        list(a = intercept,
             b = slope,
             r2 = rsq   ))))
   # Draw graph
   title <- paste(prmut)
   g1 <- ggplot(prpbmuts[prpbmuts$type == prmut,], aes(x = wk, y = num)) 
   g1 <- g1 + ggtitle(title)
   g1 <- g1 + xlab("Week") + ylab("Patients with Mutation")
   g1 <- g1 + geom_point()
   g1 <- g1 + geom_smooth(method = "lm")
   g1 <- g1 + scale_y_continuous(breaks = seq(from = 0, to = 40, by = 5))
   g1 <- g1 + scale_x_continuous(breaks = c(4, 8, 12)) 
   g1 <- g1 + annotate("text", x = 8, y = 2, label = eqn, parse = TRUE)
}
gr77 <- drawPRmut("pr77")
gr88 <- drawPRmut("pr88")
fig7788 <- ggarrange(gr77, gr88, ncol = 1, nrow = 2, align = "v")
fig7788

```


## File Information

  * File creation date: `r Sys.Date()`
  * `r R.version.string`
  * Additional session information
  
```{r echo=FALSE}
devtools::session_info()
```
  