---
title: "November Analysis"
author: "James Hunter"
date: "11 November 2016"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Packages
```{r loadpack}
suppressMessages(library(tidyverse))
suppressPackageStartupMessages(library(DescTools))
suppressPackageStartupMessages(library(caret))
suppressPackageStartupMessages(library(RColorBrewer))
suppressPackageStartupMessages(library(rafalib))
suppressPackageStartupMessages(library(WVPlots))
hmcol <- colorRampPalette(rev(brewer.pal(4, "Dark2")))(4)
options(scipen = 1000)
```

## Load Data
```{r loaddata, eval = FALSE}
load("pacdat.RData")
# Load data from SPSS file - data does not contain the variables required
# library(foreign)
# spss_data <- read.spss("tese_PBMC_final n=35.sav", to.data.frame = TRUE)
# colnames(spss_data)
# save(spss_data, file = "spss_data.RData")
act_data <- read_csv("STI_raw_activation.csv")
# check data structure
str(act_data)


## Create calculated form of variables to control rounding errors in 
## original spreadsheet (esp. csv form)
## cd4_rel and cd8_rel not used so not calculated

# log cv
act_data <- act_data %>% mutate(logcalc = log10(cviral))
cviral0 <- which(act_data$cviral == 0)
# change the logcalc that would be -inf (i.e., cviral = 0) to 0
act_data$logcalc[cviral0] <- 0

# %NK
act_data <- act_data %>% mutate(nk_pct_calc = (R3_1 + R3_2)/2) 

# cd38_index
act_data <- act_data %>% mutate(cd38_ind_calc = cd38/isto_FITC)

# hladr_index
act_data <- act_data %>% mutate(hladr_ind_calc = hla_dr/isto_PE)

# cd8_cd38_pct
act_data <- act_data %>% mutate(cd838_pct_calc = cd8_ur + cd8_lr)

# cd8_hladr_pct 
act_data <- act_data %>% mutate(cd8hla_pct_calc = cd8_ur + cd8_ul)

# cd8_cd38_hladr_pct
act_data <- act_data %>% mutate(cd838hlad_pct_calc = cd8_ur)

## Create data frame relatings pacdata$pac (paciente) to act_data$subject 

#   *  since the latter uses a different sequence

#   *  Necessary to use a data point in common, e.g., cv at 1st visit
   
# Start by creating a version of act_data with only 1st visits

act_visit12 <- act_data %>% filter(visit == 1) %>% select(subject, cviral)
# add cviral value for 2nd visit as a confirmation
act_visit12$cviral2 <- act_data$cviral[act_data$visit == 2]


#Test for concordance between the DH.xls (act_data) data frame and the original #pac_data data frame

# create empty data frame
N <- nrow(act_visit12)
concord <- data.frame(subject = integer(N),
                      pac_no = integer(N),
                      group = integer(N),
                      cviral = integer(N),
                      cv1_pac = integer(N),
                      cviral2 = integer(N),
                      cv2_pac = integer(N),
                      match = logical(N))
for (j in 1:nrow(actvisitx)) {
   concord$subject[j] <- actvisitx$subject[j]
   concord$pac_no[j] <- actvisitx$pac_no[j]
   concord$cviral[j] <- actvisitx$cviral[j]
   concord$cv1_pac[j] <- pacdata$cv1[which(pacdata$pac == 
                                              as.numeric(actvisitx$pac_no[j]))]
   concord$cviral2[j] <- actvisitx$cviral2[j]
   concord$cv2_pac[j] <- pacdata$cv2[which(pacdata$pac == 
                                              as.numeric(actvisitx$pac_no[j]))]
   concord$match[j] <- ifelse(concord$cviral[j] == concord$cv1_pac[j] &&
                                 concord$cviral2[j] == concord$cv2_pac[j], TRUE, FALSE)
# If there are NA's in the match field, change them to FALSE
   concord$match[j] <- ifelse(is.na(concord$match[j]), FALSE, concord$match[j])
   concord$group[j] <- pacdata$grupo[which(pacdata$pac == 
                                              as.numeric(actvisitx$pac_no[j]))] 
}


## There are a total of 16 records that meet the criteria. Set up a data frame
## with those 16 records and then use that to subset the act_data data frame.

match_data <- act_data[act_data$subject %in% concord$subject,]
match_data <- cbind(match_data, group = )
match_data <- match_data %>% select(subject:visit, logcalc, cd4_total, cd8_total,                                     cd38, hla_dr)
## expand the group variable by 4 visits each and 
## insert into match_data data frame
match_data$group <- rep(concord$group, each = 4)
save(match_data, act_data, file = "activation.RData")
```

## Load revised data
```{r loadrevdata}
load("activation.RData")

```


## Descriptive Stats of hla and cd38

```{r descstats1}
# First with matching data only so it can be associated with the groups
Desc(match_data$cd38, main = "cd38", plotit = TRUE)
Desc(match_data$hla_dr, main = "hla-dr", plotit = TRUE)
Desc(logcalc ~ factor(visit), data = match_data, plotit = TRUE)
Desc(logcalc ~ factor(group), data = match_data, plotit = TRUE)
Desc(cd4_total ~ factor(visit), data = match_data, plotit = TRUE)
Desc(cd4_total ~ factor(group), data = match_data, plotit = TRUE)
Desc(cd38 ~ factor(visit), data = match_data, main = "cd38 by visit", plotit = TRUE)
Desc(hla_dr ~ factor(visit), data = match_data, main = "hla_dr by visit", plotit = TRUE)
Desc(cd38 ~ factor(group), data = match_data, main = "cd38 by group", plotit = TRUE)
Desc(hla_dr ~ factor(group), data = match_data, main = "hla_dr by group", plotit = TRUE)

## Now, using act_data, therefore no grouping
Desc(act_data$cd38, main = "cd38", plotit = TRUE)
Desc(act_data$hla_dr, main = "hla-dr", plotit = TRUE)
Desc(logcalc ~ factor(visit), data = act_data, plotit = TRUE)
Desc(cd4_total ~ factor(visit), data = act_data, plotit = TRUE)
Desc(cd38 ~ factor(visit), data = act_data, main = "cd38 by visit", plotit = TRUE)
Desc(hla_dr ~ factor(visit), data = act_data, main = "hla_dr by visit", plotit = TRUE)

```

## Relation of VL and cd4 and cd8 to HLA-DR and cd38
Using test of the Ruggiero (2015) article, which stated that integrated HIV-1 DNA load was 0.5 log10 copies greater for each 50% increase in frequency of HLA-DR


## Percentage change function - deltpct
```{r deltpct}
deltpct <- function(old, new) {
   x <- (new - old) / old
   return(x)
}

```


## Do hla and cd38 increase as cd4 decreases?
Compare & graph hla and cd38 against cd4 and cd8

```{r compare}
markerset <- select(act_data, cd4_total, logcalc, 
                              hla_dr, cd4_total, cd8_total, cd38)
car::scatterplotMatrix(markerset)
cor(markerset, use = "pairwise.complete.obs", method = "spearman")
# correlation tests on cd38 against vl and cd4
cor.test(markerset$logcalc, markerset$cd38, use = "pairwise.complete.obs", method = "spearman")
cor.test(markerset$cd4_total, markerset$cd38, use = "pairwise.complete.obs", method = "spearman")
# correlation tests on hla against vl and cd4 - not expected to be significant
cor.test(markerset$logcalc, markerset$hla_dr, use = "pairwise.complete.obs", method = "spearman")
cor.test(markerset$cd4_total, markerset$hla_dr, use = "pairwise.complete.obs", method = "spearman")

## graphs showing relationship between cd38 and the two variables
# Histogram of cd38
hist38 <- ggplot(act_data, aes(x = cd38)) + geom_histogram(bins = 15, fill = "black")
hist38 <- hist38 + ggtitle("Distribution of cd38 across All Visits") 
hist38 <- hist38 + xlab("cd 38 Count") + ylab("Occurrences")
hist38
# cd38 histogram per visit
hist38v <- ggplot(act_data, aes(x = cd38)) + geom_histogram(bins = 15, fill = "black")
hist38v <- hist38v + ggtitle("Distribution of cd38 by Visit") 
hist38v <- hist38v + facet_grid(visit ~ .)
hist38v <- hist38v + xlab("cd 38 Count") + ylab("Occurrences")
hist38v

# cd38 vs vl scatterplot no visit
scat38vl <- ggplot(act_data, aes(x = cd38, y = logcalc))
scat38vl <- scat38vl + ggtitle("Distribution of cd38 and Log10 Viral Load")
scat38vl <- scat38vl + xlab("cd38") + ylab("log10 Viral Load") + labs(colour = "Visit")
scat38vl <- scat38vl 
scat38vl <- scat38vl + geom_point(shape = 16) + geom_smooth(method = "lm", na.rm = TRUE, colour = "darkred")
scat38vl

cd38vlmodel <- lm(cd38 ~ logcalc, data = act_data)
summary(cd38vlmodel)

# cd38 vs cd4 scatterplot no visit
scat38cd <- ggplot(act_data, aes(y = cd38, x = cd4_total))
scat38cd <- scat38cd + ggtitle("Distribution of cd38 and cd4 Count")
scat38cd <- scat38cd + ylab("cd38") + xlab("cd4 Total Count") + labs(colour = "Visit")
scat38cd <- scat38cd 
scat38cd <- scat38cd + geom_point(shape = 16) + geom_smooth(method = "lm", na.rm = TRUE, colour = "darkred")
scat38cd

cd38cd4model <- lm(cd38 ~ cd4_total, data = act_data)
summary(cd38cd4model)

# cd38 vs vl scatterplot with visit
scat38vl <- ggplot(act_data, aes(x = cd38, y = logcalc, colour = factor(visit)))
scat38vl <- scat38vl + ggtitle("Distribution of cd38 and Log10 Viral Load")
scat38vl <- scat38vl + xlab("cd38") + ylab("log10 Viral Load") + labs(colour = "Visit")
scat38vl <- scat38vl + scale_colour_manual(values = hmcol)
scat38vl <- scat38vl + geom_point(shape = 16) 
scat38vl

cd38vlvismod <- lm(cd38 ~ logcalc + visit, data = act_data)
summary(cd38vlvismod)
```



