---
title: "Viral Load STI"
author: "James Hunter"
date: "December 16, 2014"
output: pdf_document
---

Set up viral load for graphing using pacdat2

```{r set up viral load data}
# create vldata subset
vlset <- subset(pacdat2, type = "cv", select = c(pac, resist, value, wk))

# create VL mean for each patient
summ_vl <- by(vlset$value, vlset$pac, summary)
summ_vlwk <- by(vlset$value, vlset$wk, summary)
mean_vl <- by(vlset$value, vlset$pac, mean)
summary(mean_vl)
summ_vlwk
```


# Show boxplot of mean_vl
```{r mean_vl}
require(ggplot2)
vlbox <- ggplot(vlset, aes(x = factor(wk), y = value))
vlbox <- vlbox + geom_boxplot()
vlbox <- vlbox + stat_summary(fun.y="mean", geom="point", shape=23, size=3, fill="white")
vlbox <- vlbox + xlab("Week") + ylab("viral copies per ml (log scale)")
vlbox <- vlbox + scale_y_log10(breaks = 10 ^ (0:7))
vlbox <- vlbox + ggtitle("Viral Load Distribution by Week")
vlbox
```

The boxplot clearly shows the increase in median viral load across the patients as the number of weeks since the stopping of ART therapy increases. Note that the y-axis scale for this graph is logarithmic with the median varying between 100,000 and 1,000,000.

Difference between Patient Classes as Resistant and Non-Resistant
-----------------------------------------------------------------

```{r}
by(vlset$value, vlset$resist, summary)
vlrnr_mean <-  by(vlset$value, vlset$resist, mean)

vlrnr <- ggplot(vlset, aes(x = resist, y = value))
vlrnr <- vlrnr + geom_boxplot()
vlrnr <- vlrnr + stat_summary(fun.y="mean", geom="point", shape=23, size=3, fill="white")
vlrnr <- vlrnr + xlab("Resist/Non-Resistant") + ylab("viral copies per ml (log scale)")
vlrnr <- vlrnr + scale_y_log10(breaks = 10 ^ (0:7))
vlrnr <- vlrnr + ggtitle("Viral Load Distribution\nResistant vs. Non-Resistant")
vlrnr
```

Line Graph of Viral Load Resistant vs. Non-Resistant
----------------------------------------------------

```{r line}
vlrnrmean <- data.frame(resist = c("NR", "R"), 
                 val = c(mean(vlset$value[vlset$resist == "NR"]), 
                         mean(vlset$value[vlset$resist == "R"])))
vlrnrm <- ggplot(vlrnrmean, aes(x = resist, y = val, group = 1))
vlrnrm <- vlrnrm + geom_bar(stat = "identity", width = 0.5)
vlrnrm <- vlrnrm + xlab("Non-Resistant/Resistant") + ylab("mean viral copies per ml")
vlrnrm <- vlrnrm + scale_y_continuous(breaks = seq(100000, 600000, by = 100000))
vlrnrm <- vlrnrm + ggtitle("Viral Load Means\nResistant vs. Non-Resistant")
vlrnrm <- vlrnrm + geom_text(aes(label = round(val,2)), vjust = -0.5)
vlrnrm
```

Graph of Means of Resistant vs. Non-Resistant by Week
-----------------------------------------------------
```{r}
# set up database with week, resist and NA for values
vlrnrmeanwk <- data.frame(week = c(0, 0, 4, 4, 8, 8, 12, 12),
                          resist = rep(c("NR", "R"), 4), 
                          val = rep(NA, 8))

# add values to the database
for (i in 1:8) {
   wk <- vlrnrmeanwk[i,1]
   res <- vlrnrmeanwk[i,2]
   if (res == "NR") {
      vlrnrmeanwk$val[i] <- mean(vlset$value[vlset$resist == "NR" & vlset$wk == wk])
   }else {
      vlrnrmeanwk$val[i] <- mean(vlset$value[vlset$resist == "R" & vlset$wk == wk])   }
}
# Build plot                         
vlrnrmwk <- ggplot(vlrnrmeanwk, aes(x = week, y = val, 
                  linetype = resist, colour = resist))
vlrnrmwk <- vlrnrmwk + scale_colour_brewer(palette="Set1")
vlrnrmwk <- vlrnrmwk + geom_line(size = 2) + geom_point(size = 4,
                       colour = "black", fill = "black")
vlrnrmwk <- vlrnrmwk + xlab("Week") + ylab("mean viral copies per ml")
vlrnrmwk <- vlrnrmwk + scale_x_continuous(breaks = c(0, 4, 8, 12))
vlrnrmwk <- vlrnrmwk + ggtitle("Viral Load Means by Week\nResistant vs. Non-Resistant")
vlrnrmwk
```

