---
title: "TR Plasma Mutations"
author: "James Hunter"
date: "March 2, 2016"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Track RT Plasma Mutations by Week
---------------------------------------
```{r}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(readr))
suppressPackageStartupMessages(library(ggplot2))
# read in data from csv file
trplas <- read_csv("trplasma.csv")
# Get all the trplas mutations
trplasmutall <- colSums(trplas[-1])
wkdesigpl <- c("bl", "_12")
wkdesigpln <- length(wkdesigpl)
trmutnames <- c("tr41", "tr44", "tr65", "tr67", "tr69", "tr70", "tr74", "tr75",
              "tr103", "tr106", "tr108", "tr118", "tr179", "tr181", "tr184",
              "tr188", "tr190", "tr208", "tr210", "tr211", "tr214", "tr215",
              "tr219")
trmutnum <- length(trmutnames)
```

Create data frame repeating sequence of mutations and weeks
-----------------------------------------------------------

```{r dataframe}

trplmut <- data.frame(week = rep(wkdesigpl, each = trmutnum),
                      mutation = rep(trmutnames, times = wkdesigpln),
                      value = trplasmutall,
                      stringsAsFactors = FALSE)
# set up week factor
trplmut$week <- factor(trplmut$week, levels = wkdesigpl, 
                       labels = c("baseline", "12"))
trplmut$wknum <- rep(c(0, 12), each = trmutnum)
#set up mutation factor
trplmut$mutation <- factor(trplmut$mutation)
save(trplmut, file = "trplmut.RData")
```

Load protease database

```{r}
load("trpbmuts.RData")
```

Load ggplot2 library and plot progress -- all cases
------------------------------------------------------

The dots represent the data points themselves, the line a regression of the number of patients against time and the grey band, confidence interval. Inclusion in the confidence interval indicates that the regression line describes the data.

```{r regressfunct}
regressgraph <- function(mut){
   fit <- lm(value ~ wknum, data = trplmut[trplmut$mutation == mut,])
   eqn <- as.character(as.expression(
      substitute(italic(y) == a + b * italic(x) ,
                 list(a = format(coef(fit)[1], digits=3),
                      b = format(coef(fit)[2], digits=3)
                 ))))
   # Draw graph
   require(ggplot2)
   title <- paste0("Evolution of RT Plasma Mutation at ", mut)
   g1 <- ggplot(trplmut[trplmut$mutation == mut,], aes(x = wknum, y = value)) 
   g1 <- g1 + ggtitle(title)
   g1 <- g1 + xlab("Week") + ylab("Patients with Mutation")
   g1 <- g1 + geom_point()
   g1 <- g1 + geom_smooth(method = "lm")
   g1 <- g1 + scale_y_continuous(breaks = seq(from = 0, to = 40, by = 5))
   g1 <- g1 + scale_x_continuous(breaks = c(4, 8, 12)) 
   g1 <- g1 + annotate("text", x = 3, y = 1, label = eqn, parse = TRUE)
   print(g1)
# Determine number of weeks to achieve 0 patients with mutation
# parameters
intercept <- fit$coef[1]
slope <- fit$coef[2]
# algebraic solution
wks <- (-intercept/slope)
print(fit)
print(paste(prettyNum(wks, digits = 4), "weeks to 0 patients with mutation"))
}
```

## Print Regression Graphs

```{r printgraph}
regressgraph("tr41")
regressgraph("tr215")
regressgraph("tr190")
```

