---
title: "Calculations with Revised Reversion Groups"
author: "James R. Hunter"
date: "1 June 2018"
output:
  pdf_document: default
  word_document: default
  html_notebook: default
tables: yes
editor_options:
  chunk_output_type: inline
version: 2
---

Last week, I determined that some of the original assignments to the reversion groups were wrong. Here, **I will re-establish the groups based on the decision rule outlined in the "Total Reversion Groups" report and re-run the analyses that used the reversion groups.**

## Load Packages

```{r loadmods, echo = TRUE, message = FALSE}
library(tidyverse)
library(DescTools)
library(knitr)
library(RColorBrewer)
library(readxl)
library(ggpubr)
library(kableExtra)
library(stringr)
options(scipen = 5)
```

## Load Data

```{r loaddata}

load("pacdat.RData")
rev_groups <- readRDS("rev_groups.rds")
```

## Create new version of pacdata and pacdat2 with revised groups

```{r rev_pacdat}
# wide format
pacdata_rev <- pacdata %>% 
   mutate(grupo = as.integer(rev_groups$rev_group)) %>% 
   mutate(resist = ifelse(grupo == 0, "NR", "R"))

# Change the 9's (non-resistant controls) to 1's for patients 7 and 37 who moved from 
# controls to nones
for (i in 1:6) {
   pacdata_rev[pacdata_rev$pac == 7, 3+i] <-  1
   pacdata_rev[pacdata_rev$pac == 37, 3+i] <-  1
   pacdata_rev[pacdata_rev$pac == 11, 3+i]  <-  9 # Do the reverse for case 11 to control
}

saveRDS(pacdata_rev, file = "pacdata_rev.rds")

# long format
# cv
cv <- pacdata_rev %>% 
   select(c(pac:grupo, cv1:cv4, resist)) %>% 
   gather(type, value, cv1:cv4) %>% 
   mutate(type = "cv") %>% 
   mutate(wk = rep(c(0, 4, 8, 12), each = 38))

# cd4
cd4 <- pacdata_rev %>% 
   select(c(pac:grupo, CD41:CD412, resist)) %>% 
   gather(type, value, CD41:CD412) %>% 
   mutate(type = "cd4") %>% 
   mutate(wk = rep(c(0, 4, 8, 12), each = 38))

# cd8
cd8 <- pacdata_rev %>% 
   select(c(pac:grupo, CD81:CD812, resist)) %>% 
   gather(type, value, CD81:CD812) %>%  
   mutate(type = "cd8") %>% 
   mutate(wk = rep(c(0, 4, 8, 12), each = 38))

pacdat2_rev <- bind_rows(cv, cd4, cd8) %>% 
   mutate(wk = factor(wk)) %>% 
   mutate(type = factor(type, levels = c("cv", "cd4", "cd8")))
saveRDS(pacdat2_rev, file = "pacdat2_rev.rds")
# clean up
rm(cv, cd4, cd8)
```

## Basic Exploratory Data on Revised Groups

```{r eda_revgroup}
kable(table(pacdata_rev$grupo), caption = "Revised Group Totals", col.names = 
         c("Group", "Frequency"))
# Show comparative totals
comp_groups <- rev_groups %>% 
   select(group, rev_group) %>% 
   mutate(group = case_when(
      group == 0 ~ "control",
      group == 1 ~ "none",
      group == 2 ~ "total",
      group == 3 ~ "partial")) %>% 
   mutate(rev_group = case_when(
      rev_group == 0 ~ "control",
      rev_group == 1 ~ "none",
      rev_group == 2 ~ "total",
      rev_group == 3 ~ "partial"))

group_tots <- tibble(name = c("control", "none", "total", "partial"),
                     Original = 0L,
                     Revised = 0L)
for (i in 1:4) {
   group_tots$Original[i] <- sum(comp_groups$group == group_tots$name[i])
   group_tots$Revised[i] <- sum(comp_groups$rev_group == group_tots$name[i])
}

kable(group_tots, caption = "Comparison of Original and Revised Groups", 
       col.names = c("Group", "Original", "Revised"))
grgroup_tots <- group_tots %>% 
   gather(stage, number, -1) %>% 
   ggbarplot(x = "name", y = "number", fill = "stage", color = "stage",
             palette = "aaas",
             position = position_dodge(0.8),
             title = "Comparison of Original and Revised Groups",
             xlab = "Group",
             ylab = "Frequency",
             label = TRUE,
             lab.size = 3,
             ggtheme = theme_gray())
ggpar(grgroup_tots, legend.title = "")

## resistance
table(pacdata_rev$resist)
```

## Table 1 -- Viral Load by Reversion Group

```{r}
avgvl_rev <- pacdat2_rev %>% 
   filter(type == "cv") %>%
   mutate(grupo = case_when(
      grupo == 0 ~ "control",
      grupo == 1 ~ "none",
      grupo == 2 ~ "total",
      grupo == 3 ~ "partial"
   )) %>% 
   mutate(res_nores = factor((ifelse(grupo == "control", "NR", "R")))) %>%  
   mutate(wk = factor(wk)) %>% 
   mutate(value = log10(value)) %>% 
   group_by(grupo, wk, res_nores) %>% 
   summarize(avg = mean(value), med = median(value)) %>% 
   ungroup()
all <- pacdat2_rev %>% 
   filter(type == "cv") %>%
   mutate(res_nores = factor((ifelse(grupo == "control", "NR", "R")))) %>%  
   filter(res_nores == "R") %>% 
   mutate(wk = factor(wk)) %>% 
   mutate(value = log10(value)) %>% 
   group_by(wk) %>% 
   summarize(avg = mean(value), med = median(value))
avgvl_rev <- avgvl_rev %>% 
   add_row(grupo = "all_resist", 
           wk = all$wk,
           res_nores = "R",
           avg = all$avg, 
           med = all$med)
avgvl_rev
saveRDS(avgvl_rev, file = "avgvl_rev_fig1.rds")
# p-values by t.test
options(scipen = 1)
t.test(avgvl_rev$avg[avgvl_rev$grupo == "none"])
t.test(avgvl_rev$avg[avgvl_rev$grupo == "partial"])
t.test(avgvl_rev$avg[avgvl_rev$grupo == "total"])
```

```{r fig1draw}
# Draw Figure 1 graph -- Viral Load
cvfig1_rev <- avgvl_rev %>% 
   filter(res_nores == "R") %>% 
   ggline(x = "wk", y = "avg", 
       group = "grupo", 
       palette = "aaas", 
       color = "grupo",
       plot_type = "b", 
       legend = "bottom",
       legend.title = "",
       title = "Mean Viral Load Evolution by Reversion Group", 
       font.main = c(16),
       xlab = "Week/Test", 
       ylab = "Mean Viral Load (log10 copies/mL)", 
       ggtheme = theme_gray())
cvfig1_rev
```

### t-test of significance between week 0 and week 12 for groups

```{r cvttest}
beginendvl_rev <- pacdat2_rev %>% 
   filter(type == "cv" & wk %in% c(0, 12)) %>% 
   mutate(grupo = case_when(
      grupo == 0 ~ "control",
      grupo == 1 ~ "none",
      grupo == 2 ~ "total",
      grupo == 3 ~ "partial"
   )) %>% 
   mutate(value = log10(value))

# Control
wilcox.test(value ~ wk, data = beginendvl_rev, subset = grupo == "control")
# No reversion
wilcox.test(value ~ wk, data = beginendvl_rev, subset = grupo == "none")
# total reversion
wilcox.test(value ~ wk, data = beginendvl_rev, subset = grupo == "total")
# partial reversion
wilcox.test(value ~ wk, data = beginendvl_rev, subset = grupo == "partial")

Desc(value ~ wk, data = beginendvl_rev, plotit = TRUE)
Desc(log10(value) ~ factor(wk), data = pacdat2_rev, subset = type == "cv")

# Increase in CV for Total group 

totcvwk <- avgvl_rev %>% 
   filter(grupo == "total") %>% 
   select(wk, avg)

diff48 <- totcvwk$avg[totcvwk$wk == "8"] - totcvwk$avg[totcvwk$wk == "4"]
diff812 <- totcvwk$avg[totcvwk$wk == "12"] - totcvwk$avg[totcvwk$wk == "8"] 
print(paste("Difference wk 4 - wk 8:", diff48))
print(paste("Difference wk 8 - wk 12:", diff812))
print(paste("% Diff wk 4 - wk 8:", 100 * diff48/totcvwk$avg[totcvwk$wk == "4"]))
print(paste("% Diff wk 8 - wk 12:", 100 * diff812/totcvwk$avg[totcvwk$wk == "8"]))
diff12parttot <- avgvl_rev$avg[avgvl_rev$grupo == "total" & 
                                    avgvl_rev$wk == "12"] -
                 avgvl_rev$avg[avgvl_rev$grupo == "partial" & 
                                    avgvl_rev$wk == "12"]
diff12parttot
```

## Table 2 -- Means for CD4+ T Cells for Reversion Groups (Cells/mL)

```{r}
avgcd4_rev <- pacdat2_rev %>% 
   filter(type == "cd4") %>%
   mutate(grupo = case_when(
      grupo == 0 ~ "control",
      grupo == 1 ~ "none",
      grupo == 2 ~ "total",
      grupo == 3 ~ "partial"
   )) %>% 
   mutate(resist = factor(resist)) %>%  
   group_by(grupo, wk, resist) %>% 
   summarize(avg = mean(value), med = median(value)) %>% 
   ungroup()
all <- pacdat2_rev %>% 
   filter(type == "cd4") %>%
   mutate(res_nores = factor((ifelse(grupo == "control", "NR", "R")))) %>%  
   filter(res_nores == "R") %>% 
   mutate(wk = factor(wk)) %>% 
   group_by(wk) %>% 
   summarize(avg = mean(value), med = median(value))
avgcd4_rev <- avgcd4_rev %>% 
   add_row(grupo = "all_resist", 
           wk = all$wk,
           resist = "R",
           avg = all$avg, 
           med = all$med)

avgcd4_rev
saveRDS(avgcd4_rev, file = "avgcd4_rev_fig2.rds")
```


```{r fig2draw}
# Draw Figure 1 graph -- Viral Load
cdfig2_rev <- avgcd4_rev %>% 
   filter(resist == "R") %>% 
   ggline(x = "wk", y = "avg", 
       group = "grupo", 
       palette = "npg", 
       color = "grupo",
       plot_type = "b", 
       legend = "bottom",
       legend.title = "",
       title = "Mean CD4+ T Cell Evolution by Reversion Group", 
       font.main = c(16),
       xlab = "Week/Test", 
       ylab = "Mean CD4 T cell count (cells/mL)", 
       ggtheme = theme_gray())
cdfig2_rev
```

### t-test of Difference in CD4 resist/non-resist patients

```{r cd4ttest}
resistpacscd4_rev <- pacdat2_rev %>% 
   filter(type == "cd4" & resist == "R")
t.test(resistpacscd4_rev$value[resistpacscd4_rev$wk == 0],
       resistpacscd4_rev$value[resistpacscd4_rev$wk == 12])
```

## CD8 - Supplementary Materials

### Table S2

```{r tables2rev}
avgcd8_rev <- pacdat2_rev %>% 
   filter(type == "cd8") %>%
   mutate(grupo = case_when(
      grupo == 0 ~ "control",
      grupo == 1 ~ "none",
      grupo == 2 ~ "total",
      grupo == 3 ~ "partial"
   )) %>% 
   mutate(resist = factor(resist)) %>%  
   group_by(grupo, wk, resist) %>% 
   summarize(avg = mean(value), med = median(value))
avgcd8_rev
saveRDS(avgcd8_rev, file = "avgcd8_rev_figS1.rds")
```

### Figure S1

```{r figS1}
cdfigS1_rev <- ggline(data = avgcd8_rev, x = "wk", y = "avg", 
       group = "grupo", 
       palette = "aaas", 
       color = "grupo",
       plot_type = "b", 
       legend = "bottom",
       legend.title = "",
       title = "Mean CD8+ T Cell Evolution by Reversion Group", 
       font.main = c(16),
       xlab = "Week/Test", 
       ylab = "Mean CD8+ T cell count (cells/mL)", 
       ggtheme = theme_gray())
cdfigS1_rev
```

### CD8 Movement

```{r cd8calcs}
diffcd8_rev <- avgcd8_rev %>% 
   ungroup() %>% 
   filter(wk %in% c("0", "12")) %>% 
   mutate(wk = as.numeric(as.character(wk))) %>% 
   select(grupo, wk, avg) %>% 
   spread(grupo, value = avg) 

diffcd8_rev <- as.data.frame(t(diffcd8_rev))
colnames(diffcd8_rev) = c("wk0", "wk12")
diffcd8_rev <- diffcd8_rev %>% 
   slice(2:5) %>% 
   mutate(diff = wk12 - wk0)
Desc(diffcd8_rev$diff, plotit = FALSE)
```

## Table S3: Frequency of Resistance in PR and RT Regions (PBMC)

```{r resistfreq}
resisttot <- sum(pacdata_rev$resist == "R")
resistgrouptot <- pacdata_rev %>% 
   filter(resist == "R") %>% 
   group_by(grupo) %>% 
   summarize(tot = n()) %>% 
   mutate(pcttot = 100 * tot/resisttot)
resistgrouptot

# sum up protease post and sum of trnpos and trnnpos for resistant groups 
count_resist_group <- pacdata_rev %>% 
   filter(resist == "R") %>% 
   select(c(pac:trnnpos, resist), -c(prpre, trnpre, trnnpre,discord)) %>% 
   mutate(alltr = case_when(
      trnpos == 0 & trnnpos == 0 ~ 0,
      trnpos > 1 | trnnpos > 1 ~ 2,
      trnpos == 1 & trnnpos == 1 ~ 1
   ))
# none group (1)
nonepr <- sum(count_resist_group$prpos[count_resist_group$grupo == 1])
print(paste("Protease/None =", nonepr))
nonetr <- sum(count_resist_group$alltr[count_resist_group$grupo == 1])
print(paste("TR/None =", nonetr))

# Total Group (2)
totpr <- sum(count_resist_group$prpos[count_resist_group$grupo == 2])
print(paste("Protease/Total =", totpr))
tottrtot <- sum(count_resist_group$alltr[count_resist_group$grupo == 2] == 0)
print(paste("TR/Total Total =", tottrtot))
tottrnone <- sum(count_resist_group$alltr[count_resist_group$grupo == 2] == 1)
print(paste("TR/Total None =", tottrnone))
tottrpart <- sum(count_resist_group$alltr[count_resist_group$grupo == 2] == 2)
print(paste("TR/Total Partial =", tottrpart))

# Partial Group
partprtot <- sum(count_resist_group$prpos[count_resist_group$grupo == 3] == 0)
print(paste("Protease/Partial Total =", partprtot))
partprnone <- sum(count_resist_group$prpos[count_resist_group$grupo == 3] == 1)
print(paste("Protease/Partial None =", partprnone))
parttrtot <- sum(count_resist_group$alltr[count_resist_group$grupo == 3] == 0)
print(paste("TR/Partial Total =", parttrtot))
parttrnone <- sum(count_resist_group$alltr[count_resist_group$grupo == 3] == 1)
print(paste("TR/Partial None =", parttrnone))
parttrpart <- sum(count_resist_group$alltr[count_resist_group$grupo == 3] == 2)
print(paste("TR/Partial Partial =", parttrpart))
100 *19/resisttot
100 * 17/resisttot
100 * 13/resisttot
100 * 5/resisttot
100 * 18/resisttot
```

## Reversion Groups by Fraction and Region
Work with resistant cases only (rev_group = 1 , 2, or 3
)
```{r revgrp_frac_reg}

```

