---
title: "Revised_Mutations"
author: "James R. Hunter"
date: "5/6/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Modules

```{r loadmods, echo = TRUE, message = FALSE}
library(tidyverse)
library(DescTools)
library(knitr)
library(RColorBrewer)
library(readxl)
library(ggpubr)
library(kableExtra)
library(stringr)
options(scipen = 5)
load("prpbm2.RData")
load("trpbm2.RData")
rev_groups <- readRDS("rev_groups.rds")
```

## ID of resistant patients

```{r resist_patients}
resist_pac <- rev_groups %>% 
   filter(rev_group != 0) %>% 
   select(n)
resist_pac <- resist_pac$n
```

## PR Mutations -- Establish new mutations data frame

```{r pr_mutations}
prpbmuts <- prpbm2 %>% 
   filter(n %in% resist_pac) %>% 
   filter(value > 0) %>% 
   group_by(wk, type) %>% 
   summarize(num = n())

prpbmutstot  <- prpbmuts %>% 
   group_by(type) %>% 
   summarize(n = n())
kable(prpbmutstot)

## eliminate from consideration codons with only 1 value
not_short <- prpbmutstot %>% 
   filter(n > 1)

prpbmuts <- prpbmuts %>% 
   filter(type %in% not_short$type) 

```

# Calculate regressions and weeks to 0 for PR

```{r pr_wksto0, warning = FALSE}
wksto0 <- tibble(Mutation = NA, Weeks = NA, int = NA, beta = NA, r2 = NA)
prmutnames <- unique(prpbmuts$type)
# loop over mutations
for (i in 1:length(prmutnames)){
# Calculate regression equation
   fit <- lm(num ~ wk, data = prpbmuts[prpbmuts$type == prmutnames[i],])
# Put week to 0 patients calculation in wk0 data frame
   intercept <- fit$coef[1]
   slope <- fit$coef[2]
# algebraic solution
   wks <- (-intercept/slope)
   wksto0 <- rbind(wksto0, data_frame(Mutation = as.character(prmutnames[i]), 
                                      Weeks = wks,
                                      int = intercept,
                                      beta = slope,
                                      r2 = summary(fit)$r.squared))
}
```
## Table of Weeks to 0 Patients with Mutation

```{r tableto0}
wksto0 <- wksto0[-1,] # remove the first row with the NA values
wksto0$Weeks <- round(wksto0$Weeks, digits = 3)
wksto0$r2 <- round(wksto0$r2, digits = 3)
rownames(wksto0) <- NULL
kable(wksto0)
```
The next table shows the PR mutations that are either infinite (because the values are all the same) or will never get to 0 (Weeks are negative.). I will eliminate these from further analysis

```{r remove_cases_pr}
wksto0 %>% 
   filter(Weeks < 0 | Weeks == Inf) 
edit_out_pr <- wksto0$Mutation[wksto0$Weeks == Inf | wksto0$Weeks < 0] 
   
#Remove these codons from the analysis
prpbmuts <- prpbmuts %>% filter(!(type %in% edit_out_pr))
prmutnames <- unique(prpbmuts$type)
# Reinitialize wksto0
wksto0 <- data_frame(Mutation = NA, Weeks = NA, int = NA, beta = NA, r2 = NA)
# Recalculate regression
# loop over mutations
for (i in 1:length(prmutnames)){
# Calculate regression equation
   fit <- lm(num ~ wk, data = prpbmuts[prpbmuts$type == prmutnames[i],])
# Put week to 0 patients calculation in wk0 data frame
   intercept <- fit$coef[1]
   slope <- fit$coef[2]
# algebraic solution
   wks <- (-intercept/slope)
   wksto0 <- rbind(wksto0, data_frame(Mutation = as.character(prmutnames[i]), 
                                      Weeks = wks,
                                      int = intercept,
                                      beta = slope,
                                      r2 = summary(fit)$r.squared))
}
prwksto0 <- wksto0[-1,]
```

## TR Mutations -- Establish new mutations data frame

```{r tr_mutations}
trpbmuts <- trpbm2 %>% 
   filter(n %in% resist_pac) %>% 
   filter(value > 0) %>% 
   group_by(wk, type) %>% 
   summarize(num = n())
trpbmutstot  <- trpbmuts %>% 
   group_by(type) %>% 
   summarize(n = n())
kable(trpbmutstot)

## eliminate from consideration codons with only 1 value
not_short <- trpbmutstot %>% 
   filter(n > 1)

trpbmuts <- trpbmuts %>% 
   filter(type %in% not_short$type) 

```

# Calculate regressions and weeks to 0 for TR

```{r tr_wksto0, warning = FALSE}
wksto0 <- tibble(Mutation = NA, Weeks = NA, int = NA, beta = NA, r2 = NA)
trmutnames <- unique(trpbmuts$type)
# loop over mutations
for (i in 1:length(trmutnames)){
# Calculate regression equation
   fit <- lm(num ~ wk, data = trpbmuts[trpbmuts$type == trmutnames[i],])
# Put week to 0 patients calculation in wk0 data frame
   intercept <- fit$coef[1]
   slope <- fit$coef[2]
# algebraic solution
   wks <- (-intercept/slope)
   wksto0 <- rbind(wksto0, data_frame(Mutation = as.character(trmutnames[i]), 
                                      Weeks = wks,
                                      int = intercept,
                                      beta = slope,
                                      r2 = summary(fit)$r.squared))
}
```
## Table of Weeks to 0 Patients with Mutation

```{r tableto0a}
wksto0 <- wksto0[-1,] # remove the first row with the NA values
wksto0$Weeks <- round(wksto0$Weeks, digits = 3)
wksto0$r2 <- round(wksto0$r2, digits = 3)
rownames(wksto0) <- NULL
kable(wksto0)
```

```{r remove_cases_tr}
wksto0 %>% 
   filter(Weeks < 0 | Weeks == Inf) %>% 
   kable()
edit_out_tr <- wksto0$Mutation[wksto0$Weeks == Inf | wksto0$Weeks < 0] 
   
#Remove these codons from the analysis
trpbmuts <- trpbmuts %>% 
   filter(!(type %in% edit_out_tr)) %>% 
   ungroup()
trmutnames <- unique(trpbmuts$type)
# Reinitialize wksto0
wksto0 <- data_frame(Mutation = NA, Weeks = NA, int = NA, beta = NA, r2 = NA)
# Recalculate regression
# loop over mutations
for (i in 1:length(trmutnames)){
# Calculate regression equation
   fit <- lm(num ~ wk, data = trpbmuts[trpbmuts$type == trmutnames[i],])
# Put week to 0 patients calculation in wk0 data frame
   intercept <- fit$coef[1]
   slope <- fit$coef[2]
# algebraic solution
   wks <- (-intercept/slope)
   wksto0 <- rbind(wksto0, data_frame(Mutation = as.character(trmutnames[i]), 
                                      Weeks = wks,
                                      int = intercept,
                                      beta = slope,
                                      r2 = summary(fit)$r.squared))
}
trwksto0 <- wksto0[-1,]
```

## Save new files

```{r save_rev_files}
saveRDS(trpbmuts, file = "trpbmuts_rev.rds")
saveRDS(prpbmuts, file = "prpbmuts_rev.rds")
saveRDS(trwksto0, file = "trweeksto0.rds")
```


## Table 8 Mean Weeks until Complete Reversion

```{r}
nrti_muts <- c("tr41",  "tr210", "tr215", "tr67",  "tr70",  "tr219", "tr184")
nnrti_muts <- c("tr103", "tr181", "tr190")
pi_muts <- c("pr10", "pr20", "pr24", "pr36", "pr46", "pr48",
             "pr53", "pr54", "pr63", "pr71", "pr73", "pr82",
             "pr84", "pr90")

prpbmuts <- readRDS("prpbmuts_rev.rds")
trpbmuts <- readRDS("trpbmuts_rev.rds")

prsumm <- prwksto0 %>% 
   summarize(mean = mean(Weeks), stddev = sd(Weeks))
prsumm

nrtisumm <- trwksto0  %>% 
   filter(Mutation %in% nrti_muts) %>% 
   summarize(mean = mean(Weeks), stddev = sd(Weeks))
nrtisumm

nnrtisumm <- trwksto0  %>% 
   filter(Mutation %in% nnrti_muts) %>% 
   summarize(mean = mean(Weeks), stddev = sd(Weeks))
nnrtisumm
```

## TAMS and 184

```{r}
tam1 <- c("tr41", "tr210", "tr215")
tam2 <- c("tr67", "tr70", "tr219")

tam1no <- trpbmuts  %>% 
   filter(type %in% tam1) %>% 
   group_by(wk) %>% 
   summarize(mean = mean(num), stddev = sd(num))
tam1no

tam1summ <- trwksto0  %>% 
   filter(Mutation %in% tam1) %>% 
   summarize(mean = mean(Weeks), stddev = sd(Weeks))
tam1summ

tam2summ <- trwksto0  %>% 
   filter(Mutation %in% tam2) %>% 
   summarize(mean = mean(Weeks), stddev = sd(Weeks))
tam2summ

tam2no <- trpbmuts  %>% 
   filter(type %in% tam2) %>% 
   group_by(wk) %>% 
   summarize(mean = mean(num), stddev = sd(num))
tam2no

alltamssumm <- trwksto0  %>% 
   filter(Mutation != "tr184") %>% 
   summarize(mean = mean(Weeks), stddev = sd(Weeks))
alltamssumm

codon184summ <- trwksto0  %>% 
    filter(Mutation == "tr184")
codon184summ

codon184 <- trpbmuts  %>% 
   filter(type == "tr184")
codon184


```



