---
title: "Table 3 Mean VL, CD4, CD8 by Group"
author: "James R. Hunter"
date: "31/7/2018"
output: pdf_document
tables: yes
editor_options:
  chunk_output_type: inline
version: 2
---

## Objective

Objective is to check and revise the p values for Table 3 (Means for Viral Load by Reversion Group)


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Packages

```{r loadmods, echo = TRUE, message = FALSE}
library(tidyverse)
library(DescTools)
library(knitr)
library(RColorBrewer)
library(readxl)
library(ggpubr)
library(kableExtra)
library(stringr)
options(scipen = 1)
```

## Load Data

```{r loaddata}
##avgvl_rev <- readRDS("../avgvl_rev_fig1.rds")
##kable(avgvl_rev)
# pacdata_rev <- readRDS("pacdata_rev.rds")
pacdat2_rev <- readRDS("../pacdat2_rev.rds")
```

## Wilcoxon tests of difference within groups between weeks [Not executed]

Do Wilcoxon tests of differences between visits within groups. However, tests do not show any difference among the groups. All are suggested to be not significant. However, ranks are the same, except for inversion of 3 and 4 for "none" class

```{r wilcox, eval = FALSE}
# p-values by t.test
options(scipen = 1)
wilcox.test(avgvl_rev$avg[avgvl_rev$grupo == "none"], alternative = "two.sided")
wilcox.test(avgvl_rev$avg[avgvl_rev$grupo == "partial"], alternative = "two.sided")
wilcox.test(avgvl_rev$avg[avgvl_rev$grupo == "total"], alternative = "two.sided")
rank(avgvl_rev$avg[avgvl_rev$grupo == "none"])
rank(avgvl_rev$avg[avgvl_rev$grupo == "partial"])
rank(avgvl_rev$avg[avgvl_rev$grupo == "total"])
```

## Table of means paired's test of Viral Load by Reversion Group

```{r}
resistpaccv_rev <- pacdat2_rev %>% 
   filter(type == "cv" & resist == "R") %>% 
   mutate(grupo = case_when(grupo == 1 ~ "None",
                            grupo == 2 ~ "Total",
                            grupo == 3 ~ "Partial")) %>% 
   mutate(logcv = log10(value))

avglogvl <- resistpaccv_rev %>% 
   group_by(grupo, wk) %>% 
   summarize(mean(logcv))

avglogvl

paired_t_test <- function(group) {
   base <- resistpaccv_rev$logcv[resistpaccv_rev$wk == 0 & 
                             resistpaccv_rev$grupo == group]
   end <- resistpaccv_rev$logcv[resistpaccv_rev$wk == 12 & 
                             resistpaccv_rev$grupo == group]
   t.test(end, base, paired = TRUE)
}
# None
paired_t_test("None")
# Partial
paired_t_test("Partial")
#Total
paired_t_test("Total")

## All resistant cases
all_base <- resistpaccv_rev$logcv[resistpaccv_rev$wk == 0]
all_end <- resistpaccv_rev$logcv[resistpaccv_rev$wk == 12]
t.test(all_end, all_base, paired = TRUE)
```

## Overall resistance t-test

```{r overall}
Desc(resistpaccv_rev$logcv)
t.test(resistpaccv_rev$logcv[resistpaccv_rev$wk == 12],
       resistpaccv_rev$logcv[resistpaccv_rev$wk == 0],
       paired = TRUE)
```

## Table of means paired t test of CD4 by Reversion Group

```{r}
resistpaccd4_rev <- pacdat2_rev %>% 
   filter(type == "cd4" & resist == "R") %>% 
   mutate(grupo = case_when(grupo == 1 ~ "None",
                            grupo == 2 ~ "Total",
                            grupo == 3 ~ "Partial")) 

avgcd4 <- resistpaccd4_rev %>% 
   group_by(grupo, wk) %>% 
   summarize(mean(value))

avgcd4

paired_t_test <- function(group) {
   base <- resistpaccd4_rev$value[resistpaccd4_rev$wk == 12 & 
                             resistpaccd4_rev$grupo == group]
   end <- resistpaccd4_rev$value[resistpaccd4_rev$wk == 0 & 
                             resistpaccd4_rev$grupo == group]
   t.test(end, base, paired = TRUE)
}
# None
paired_t_test("None")
# Partial
paired_t_test("Partial")
#Total
paired_t_test("Total")
```

## Average Decline for REsistant Patients

```{r declineCD4}
resistpaccd4_rev$decline <- 
   resistpaccd4_rev$value[resistpaccd4_rev$wk == "12"] -
   resistpaccd4_rev$value[resistpaccd4_rev$wk == "0"]
mean(resistpaccd4_rev$decline)   
```

## Overall t-test for CD4

```{r overall_cd4}
Desc(resistpaccd4_rev$value)
t.test(resistpaccd4_rev$value[resistpaccd4_rev$wk == 12],
       resistpaccd4_rev$value[resistpaccd4_rev$wk == 0],
       paired = TRUE)
wilcox.test(resistpaccd4_rev$value[resistpaccd4_rev$wk == 0],
       resistpaccd4_rev$value[resistpaccd4_rev$wk == 12],
       paired = TRUE, alternative = "two.sided")
```

# CD8

## Table of means paired t test of CD8 by Reversion Group

```{r cd8groups}
resistpaccd8_rev <- pacdat2_rev %>% 
   filter(type == "cd8" & resist == "R") %>% 
   mutate(grupo = case_when(grupo == 1 ~ "None",
                            grupo == 2 ~ "Total",
                            grupo == 3 ~ "Partial")) 

avgcd8 <- resistpaccd8_rev %>% 
   group_by(grupo, wk) %>% 
   summarize(mean(value))

avgcd8

paired_t_test <- function(group) {
   base <- resistpaccd8_rev$value[resistpaccd8_rev$wk == 12 & 
                             resistpaccd8_rev$grupo == group]
   end <- resistpaccd8_rev$value[resistpaccd8_rev$wk == 0 & 
                             resistpaccd8_rev$grupo == group]
   t.test(end, base, paired = TRUE)
}
# None
paired_t_test("None")
# Partial
paired_t_test("Partial")
#Total
paired_t_test("Total")
```

## Average Decline for REsistant Patients

```{r declineCD8}
resistpaccd8_rev$change <- 
   resistpaccd8_rev$value[resistpaccd4_rev$wk == "12"] -
   resistpaccd8_rev$value[resistpaccd4_rev$wk == "0"]
mean(resistpaccd8_rev$change)
sd(resistpaccd8_rev$change)
```

## Overall t-test for CD8

```{r overall_cd8}
Desc(resistpaccd8_rev$value)
t.test(resistpaccd8_rev$value[resistpaccd8_rev$wk == 12],
       resistpaccd8_rev$value[resistpaccd8_rev$wk == 0],
       paired = TRUE)
wilcox.test(resistpaccd8_rev$value[resistpaccd8_rev$wk == 0],
       resistpaccd8_rev$value[resistpaccd8_rev$wk == 12],
       paired = TRUE, alternative = "two.sided")
```




